---
title: "STA4063 - Lab sheet - Logistic Regression"
output: pdf_document
date: "2026-06-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Learning Objectives

By the end of this lab, you will be able to:

1. Build a Bayesian logistic regression model.
2. Specify likelihood and prior distributions.
3. Fit the model using JAGS.
4. Diagnose MCMC convergence.
5. Interpret posterior distributions and odds ratios.

---



```{r setup, message=FALSE}

library(rjags)
library(coda)
library(ggplot2)
library(dplyr)
```


a) The dataset contains clinical measurements from patients.

```{r}
url <- 
"https://raw.githubusercontent.com/jbrownlee/Datasets/master/pima-indians-diabetes.data.csv"


data <- read.csv(
url,
header=FALSE
)


names(data)<-c(
"Pregnancy",
"Glucose",
"BloodPressure",
"SkinThickness",
"Insulin",
"BMI",
"DiabetesPedigree",
"Age",
"Outcome"
)


head(data)
```


b) Find the number of diabetic and non-diabetic patients.

```{r}
library(dplyr)

data %>%
  summarise(
    n_diabetic = sum(Outcome == 1),
    n_nondiabetic = sum(Outcome == 0),
    total = n()
  )
```

c) Calculate the proportion of diabetic patients.

```{r}
data %>% summarise(
  Diabetic_proprtion = sum(data$Outcome==1) / data %>% nrow() , 
  nondiabetic_proportion = sum(data$Outcome==0) / data %>% nrow()
)


```

d) Calculate mean glucose level by diabetes status and discuss the result.

```{r}
#glimpse(data)
mean(data$Glucose,data$Outcome)
```


e) Standardise the continuous predictors.

```{r}
data_Set <- data %>% 
  mutate(Glucose=scale(Glucose),
         BMI=scale(BMI),
         Age=scale(Age)
   )
```


f) Write the JAGS model specification.

```{r}
model_string23 <- "
model{

for(i in 1:n){

y[i]~ dbern(p[i])

logit(p[i]) <- beta0 + beta1*glucose[i]+ beta2*bmi[i] + beta3*age[i]

}

beta0 ~ dnorm(0,0.01)
beta1 ~ dnorm(0,0.01)
beta2 ~ dnorm(0,0.01)
beta3 ~ dnorm(0,0.01)

}
"

```

g) Create the data list required by JAGS.

```{r}
data_list<- list(
  y = data$Outcome ,
  glucose=as.vector(data_Set$Glucose),
  bmi= as.vector(data_Set$BMI),
  age =as.vector(data_Set$Age) ,
  n = nrow(data_Set)
)


```

h) Compile the JAGS model, run MCMC sampling, and obtain posterior samples.

```{r}

# Compile model
model <- jags.model(
  textConnection(model_string23),
  data     = data_list,
  n.chains = 3
)
# Run MCMC
update(model,500)
# Obtain posterior samples
samples_logit <- coda.samples(
  model= model,
  variable.names = c("beta0","beta1", "beta2","beta3"),
  n.iter = 10000
) 


```


i) Perform convergence diagnostics and discuss whether the chains have converged.

```{r}
plot(samples_logit)

```

j) Plot posterior densities of regression coefficients  and calculate posterior summaries.
```{r}
# Density plots


# posterior mean

#  posterior standard deviation


# 95% credible interval
```

k) Convert the posterior estimates into odds ratios and interpret the effect of each predictor variable
```{r}

```


